Gravitational waves from colliding black holes

### 2017-05-04 Report/Draft/draft.tex
Report latest

\begin{document}
\maketitle
\begin{abstract}
A complete understanding of the LIGO detectors' ability to accurately measure spin parameters is of key importance to understanding the astrophysics of binary black hole systems. Of particular significance is an ability to determine spin orientations as a means of distinguishing between binary formation models. In this work we overview the core physics of gravitational waves, relativistic precession in binary systems, the LIGO detectors, and the techniques involved in parameter estimation. We then examine the specific effects of precession on a gravitational waveform. The technique of 'matching' between waveforms is used as a computationally cheap way to explore degeneracies in the parameter space. A series of signals are then injected into simulated detector noise, and a full parameter estimation is performed on the data segments. Using the match results as a guide, we target these injections to areas in the parameter space where we expect to be particularly sensitive and insensitive to spin parameters. Inference for a full range of inclinations is performed, and we identify areas where spin orientations are consistently incorrectly recovered. We then briefly examine the impact the upcoming VIRGO detector will have on spin estimation.
\end{abstract}
\tableofcontents
\newpage

@@

Overall the results indicate that the current inference methods consistently underestimate values of $\chi_p$, with all but 5 out of the 20 inference runs excluding the injected value of $\chi_p$. The match results and waveform plots demonstrate that the phase of the GW signal is an important parameter in precessing systems unlike in non-precessing systems, leading to an extra degree of freedom in precessing systems. The matching process was used to highlight the way in which mass ratios, polarisations, inclinations and phases affect precessing waveforms, particularly our ability to distinguish bewteen precessing and non-precessing signals. Degenerate areas of the parameter space were also identified using this technique, and it was identified that near equal mass ratios, precessing and non-precessing waveforms are effectively entirely degenerate. It was also shown that other parameters such as source location, polarisation and inclination affect the way that the modulation effect manifests itself in a specific signal. This leads to a key observation that the strength of signal modulation in a GW waveform is not necessarily proportional to the amount of precession in the binary itself.

The inference pipeline used to perform parameter estimation was then described, and results were compared for a case of $\chi_p$ being recovered, and $\chi_p$ being excluded. Two explanations were proposed for the failure to accurately infer $\chi_p$, one being that the inference method is not working properly, and the other that there are so many intrinsic degeneracies in precessing waveforms that the inference process cannot distinguish between them. A way of distinguishing between these was also proposed, by performing a match between the recovered MAP waveform and the injected signal. This match was consistently high despite the inaccuracies in the posteriors, and so the source of the inaccuracy is likely to be intrinsic degenerecy. If these results are correct, this has significant implications for the future of spin estimation, as even improvements in detector accuracy and inference methodologies are unlikely to be able to make significant advancements in spin recovery if the signals are inherently degenerate. This would also explain the apparent lack of consistency between match results and spin inaccuracy results, as the ability to recover $\chi_p$ appears to be independent of the match between precessing and non-precessing waveforms.
\section{Reflective Statement}
I now need a therapist

>>>>

\begin{document}
\maketitle
\begin{abstract}
A complete understanding of the LIGO detectors' ability to accurately measure spin parameters is of key importance to understanding the astrophysics of binary black hole systems. Of particular significance is an ability to determine spin orientations as a means of distinguishing between binary formation models. In this work we overview the core physics of gravitational waves, relativistic precession in binary systems, the LIGO detectors, and the techniques involved in parameter estimation. We then examine the specific effects of precession on a gravitational waveform. The technique of 'matching' between waveforms is used as a computationally cheap way to explore degeneracies in the parameter space. A series of signals are then injected into simulated detector noise, and a full parameter estimation is performed on the data segments. Using the match results as a guide, we target these injections to areas in the parameter space where we expect to be particularly sensitive and insensitive to spin parameters. A total of 20 signals with $\chi_p=0.9$ across a full range of inclinations were injected and recovered, with all but 5 results having the injected value of $\chi_p$ excluded, and even in these cases the value of $\chi_p$ is consistently underestimated. Comparisons between the match and inference results suggest that matching is not a successful predictor of the inference results, and intrinsic degeneracies in the waveforms are suggested as the source of the inaccuracy. Finally we briefly examine the impact of the Virgo detector on spin estimation, and present results that indicate that the impact will be marginal.
\end{abstract}
\tableofcontents
\newpage

@@

Overall the results indicate that the current inference methods consistently underestimate values of $\chi_p$, with all but 5 out of the 20 inference runs excluding the injected value of $\chi_p$. The match results and waveform plots demonstrate that the phase of the GW signal is an important parameter in precessing systems unlike in non-precessing systems, leading to an extra degree of freedom in precessing systems. The matching process was used to highlight the way in which mass ratios, polarisations, inclinations and phases affect precessing waveforms, particularly our ability to distinguish bewteen precessing and non-precessing signals. Degenerate areas of the parameter space were also identified using this technique, and it was shown that near for equal mass ratios, precessing and non-precessing waveforms are effectively entirely degenerate. It was also shown that other parameters such as source location, polarisation and inclination affect the way that the modulation effect manifests itself in a specific signal. This leads to a key observation that the strength of signal modulation in a GW waveform is not necessarily proportional to the amount of precession in the binary itself.

The inference pipeline used to perform parameter estimation was then described, and a series of injections with high mass ratio and high $\chi_p$ were injected into detector noise. The standard inference process was performed, and results were compared for a case of $\chi_p$ being recovered comparably well, and $\chi_p$ being excluded. Neither sets of posteriors were fully accurate, but the case of $\chi_p$ being excluded actually provided more accurate results on the whole. Inference results were presented for a signal which appeared to exhibit high precession, however the MAP values on this signal still exluded the injected value of $\chi_p$ while still showing a strong match (0.96) with the injected signal. Posteriors of derived spin parameters were shown for a full range of inclinations for signals at 200Mpc and 800Mpc, and their results were compared with each other and with the predictions made by the matches. No clear correlations between any of these were observed, indicating that the relationship between matches and inference results is more complicated than previously thought. Finally a set of inference runs includingthe Virgo detector at a range of inclinations were presented and compared with previous results without Virgo. These results suggested that Virgo will not significantly improve the accuracy on $\chi_p$ due to the fact that the morphology of the signal and the amplitude modulation is the same in the Virgo detector as in H1 and L1. Although the specific set of parameters used in these injections greatly limits the generality of this result.

Two explanations were proposed for the failure to accurately infer $\chi_p$, one being that the inference method is not working properly, and the other that there are so many intrinsic degeneracies in precessing waveforms that the inference process cannot distinguish between them. A way of distinguishing between these was also proposed, by performing a match between the recovered MAP waveform and the injected signal. This match was consistently high despite the inaccuracies in the posteriors, and so the source of the inaccuracy is likely to be intrinsic degenerecy. If these results are correct, this has significant implications for the future of spin estimation, as even improvements in detector accuracy and inference methodologies are unlikely to be able to make significant advancements in spin recovery if the signals are inherently degenerate. Especially if two signals have a good match for the majority of the inspiral and early merger, if there is only significant disagreement around the end of the merger, detectors will struggle to resolve this difference due to noise. This would also explain the apparent lack of consistency between match results and spin inaccuracy results, as the ability to recover $\chi_p$ appears to be independent of the match between precessing and non-precessing waveforms.
\section{Reflective Statement}
I now need a therapist
